---
title: "RR Proj 1"
author: "CodingInR"
date: "6/12/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Code to read and prep data
```{r read}
require("dplyr")
require("ggplot2")
#Dataset URL
URL <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
#Check if zip has been downloaded. If not, download, read csv into two data frames:
#1) original 2) original with NA's omitted
if(!file.exists("data.zip")){
  download.file(URL,"data.zip")
  unzip("data.zip")
}
if(!exists("dat")){
  #dat is unchanged dataset, datNoH removes NA's
  dat <- read.csv("activity.csv", header = TRUE, sep = ",")
  datNoH <- na.omit(dat)
  }
```

### Mean number of steps per day (ignoring missing values in the dataset)
#### 1) Total number of steps per day
```{r meansteps}
#use dplyr to group by date and calculate sum of steps per day
stepsDay <- datNoH %>% group_by(date) %>% summarize_all(sum)
```
#### 2) Histogram of the total numebr of steps taken each day
```{r}
ggplot(stepsDay, aes(x=steps)) + geom_histogram()
```

#### 3) Calculate and report mean and median of the total number of steps per day

```{r}
summary(stepsDay$steps)
```
or
```{r}
meanSteps <- mean(stepsDay$steps) %>% print()
medSteps <- median(stepsDay$steps) %>% print()
```
### Average daily activity pattern
#### 1) Time series plot of interval vs average steps taken across days

```{r}
stepsint <- datNoH %>% group_by(interval) %>% summarize(MeanSteps = mean(steps))
ggplot(data = stepsint, aes(x=interval, y = MeanSteps))+geom_line(color = "blue")

```

#### 2) Which 5 minute interval contains the max number of steps?
```{r}
as.integer(stepsint[which.max(stepsint$MeanSteps),1])
```

### Imputing missing values
#### 1) Number of missing values in dataset
```{r}
length(dat$steps)-length(datNoH$steps)

```
#### 2) Fill in missing values in dataset.  Approach: if all NA's for a day, replace with zero, else, replace NA with average of steps for the day.
#### 3) Create new dataset with missing data filled in

```{r}
require("zoo")
remNA <- function(x,...) if (all(is.na(x))) 0 else mean(x,...)
datFilled <- transform(dat, steps = na.aggregate(steps, by = date, FUN = remNA))
head(datFilled)
```

#### 4) Repeat histogram and mean/media. Compare new(filled in vales) with original (missing data). Do these differ?  What's the impact of imputing?

```{r}
stepsDayFilled <- datFilled %>% group_by(date) %>% summarize_all(sum)
ggplot(stepsDayFilled, aes(x=steps)) + geom_histogram()
summary(stepsDayFilled$steps)
#show a measure of skewness between data sets
summary(stepsDay$steps) - summary(stepsDayFilled$steps)
```

##### Yes, they differ. The filled in dataset skews slightly higher overall, while having a minimum of zero and the same maximum. The impact is a loss of precision.

### Weekdays vs Weekends
#### 1) Create a new weekday/weekend factor variable in data

```{r}
datFilled$date <- as.Date(datFilled$date)
datFilled$wday <- as.factor(weekdays(datFilled$date))
table(datFilled$wday)
class(datFilled$wday)
daysweek <- c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday')
datFilled$wday <- factor(datFilled$wday %in% daysweek, levels = c(FALSE, TRUE), labels = c('weekend','weekday'))
table(datFilled$wday)

```
#### 2) Panel plot of timeseries of interval vs average number of steps taken by weekday or weekend

```{r}
require("gridExtra")
Set1 <- datFilled[datFilled$wday == 'weekend',]
Set2 <- datFilled[datFilled$wday == 'weekday',]

SetA <- Set1 %>% group_by(interval) %>% summarize(MeanSteps = mean(steps))
SetB <- Set2 %>% group_by(interval) %>% summarize(MeanSteps = mean(steps))

P1 <- ggplot(data = SetA, aes(x=interval, y = MeanSteps))+geom_line(color = "red")+ggtitle("Weekend")
P2 <- ggplot(data = SetB, aes(x=interval, y = MeanSteps))+geom_line(color = "green")+ggtitle("Weekday")

grid.arrange(P1,P2, ncol =1)
```
